# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

1- Download assessement data file if it's not existed in working directory.

2- Unzip the assessement zip file(activity.zip) into the working directory.

3- load dataset file(activity.csv) into d variable.

```{r}

if(!file.exists("activity.zip")){
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="activity.zip", method="curl")  
}

if(!file.exists("activity.csv")){
  unzip("activity.zip")
}

d <- read.csv("activity.csv", header=T)
```

## What is mean total number of steps taken per day?
```{r}

mean_steps <- c()
median_steps <- c()
taken_per_day <- c()
steps <- c()
interval <- c()
interval_mean <- c()
interval_median <- c()

for(a in names(table(d$date))){
  taken_per_day <- c(taken_per_day, a)
  steps <- c(steps, sum(d[d$date %in% c(a), "steps"], na.rm=T))
  mean_steps <- c(mean_steps, mean(d[d$date %in% c(a), "steps"], na.rm=T))
  median_steps <- c(median_steps, median(d[d$date %in% c(a), "steps"], na.rm=T))
  interval <- c(interval, sum(d[d$date %in% c(a), "interval"], na.rm=T))
  interval_mean <- c(interval_mean, mean(d[d$date %in% c(a), "interval"], na.rm=T))
  interval_median <- c(interval_median, median(d[d$date %in% c(a), "interval"], na.rm=T))
}

steps_taken_per_day <- data.frame(day=taken_per_day, steps, mean_steps, median_steps, interval, interval_mean, interval_median)

steps_taken_per_day

summary(steps_taken_per_day$steps)
summary(steps_taken_per_day$mean_steps)
summary(steps_taken_per_day$median_steps)

plot.new()
hist(steps_taken_per_day$steps , main="Steps Taken by Day", xlab="days", ylab="steps", breaks=length(steps_taken_per_day$steps), col=rainbow(length(steps_taken_per_day$steps)))
#hist(steps_taken_per_day$steps , main="Steps Taken by Day", xlab="days", ylab="steps", breaks=length(steps_taken_per_day$steps), col=rainbow(length(steps_taken_per_day$steps)), xaxt="n", yaxt="n")
#axis(side=1, at=steps_taken_per_day$steps, labels=steps_taken_per_day$day)

# axis at
#at <- c()
#for(i in 0:5){
#  at <- c(at, (round(length(steps_taken_per_day$steps)/5)*i))
#}
#labels <- round(seq(from=min(steps_taken_per_day$steps), to=max(steps_taken_per_day$steps), length.out=6))
#axis(side=2, at=at, labels=labels)
```


## What is the average daily activity pattern?
```{r}
plot(steps_taken_per_day$mean_steps, type="l", xlab="days", ylab="average daily activity")
```

## Imputing missing values
```{r}

missings_count <- nrow(d[is.na(d$steps), ])

print(paste("Total number of rows with NAs: ", missings_count))

data_with_missings_filled <- d

data_with_missings_filled[is.na(data_with_missings_filled$steps), "steps"] <- seq(missings_count)


##########################################################
mean_steps <- c()
median_steps <- c()
taken_per_day <- c()
steps <- c()
interval <- c()

for(a in names(table(data_with_missings_filled$date))){
  taken_per_day <- c(taken_per_day, a)
  steps <- c(steps, sum(data_with_missings_filled[data_with_missings_filled$date %in% c(a), "steps"], na.rm=T))
  mean_steps <- c(mean_steps, mean(data_with_missings_filled[data_with_missings_filled$date %in% c(a), "steps"], na.rm=T))
  median_steps <- c(median_steps, median(data_with_missings_filled[data_with_missings_filled$date %in% c(a), "steps"], na.rm=T))
  interval <- c(interval, sum(data_with_missings_filled[data_with_missings_filled$date %in% c(a), "interval"], na.rm=T))
}

missings_filled_per_day <- data.frame(day=taken_per_day, steps, mean_steps, median_steps, interval)

missings_filled_per_day

summary(missings_filled_per_day)

##########################################################

hist(data_with_missings_filled$steps)

```


## Are there differences in activity patterns between weekdays and weekends?
```{r}

missings_filled_per_day$weekday_weekend <- weekdays(as.Date(missings_filled_per_day$day, "%Y-%m-%d")) %in% c("Sunday")

missings_filled_per_day[missings_filled_per_day$weekday_weekend, "weekday_weekend"] <- "weekend"
missings_filled_per_day[missings_filled_per_day$weekday_weekend == F, "weekday_weekend"] <- "weekday"

missings_filled_per_day

plot.new()
plot(missings_filled_per_day[missings_filled_per_day$weekday_weekend == "weekend", "steps"], type="l", main="weekend")
plot(missings_filled_per_day[missings_filled_per_day$weekday_weekend == "weekday", "steps"], type="l", main="weekday")
```

